---
title: "Time Series Analysis for Climo with Wind Data"
output: 
  # html_document:
  rmarkdown::github_document
#
  # beamer_presentation:
  #   slide_level: 2
  #   keep_tex: TRUE
  #   fig_caption: yes
  #   toc: true

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(png)
library(grid)
library(quantreg)
library(highfrequency)
library(TTR)
library(lubridate)
library(timeSeries)
library(dplyr)
library(ggplot2)
library("gridExtra")
library("cowplot")
theme_set(theme_grey())
library(reshape2)
library(scales)
source("src/TSA_source.R")
source("TSA_Data_Source.R")
source("src/plotting_functions.R")
```


# Extraction

```{r, echo = TRUE, warning=FALSE}

C_data <- read.csv(file = 'Raw Sensor Data/Climo_co_corr_IISc_1_10.csv', header = TRUE, sep = ";")
cTime <-  as.POSIXct(C_data$Time)#(,"%Y-%m-%d %H:%M:%S", tz = "")
ws_c <- as.numeric(mean(diff(cTime)))*60 # sampling frequecy (seconds)
xts.c <- xts(x = C_data$Value, order.by = cTime)


daily_threshold <- floor(1170) # Expected number of samples per day
frequency_tolerance.c = 10
frequency_tolerance.w = 10
fault_window = "hours" # to identify faults in the data
std_tolerance = 0.01
dir.create("plots",showWarnings=FALSE,recursive=TRUE)

```



## Main dataframe

```{r overall_dataset, echo=TRUE, warning=FALSE, fig.height=5}

avg_time = 1

# df.new <- get_ts_master_dataFrame('Raw Sensor Data/Wind_direction_corr_IISc_1_20.csv', 
#                               'Raw Sensor Data/Climo_co_corr_IISc_1_10.csv', 
#                               'Raw Sensor Data/Wind_speed_corr_IISc_1_20.csv', 
#                               avg_time)
df.new <- getMasterdf_September(avg_time)
refTime <- as.POSIXct(df.new$time)

# Plot overall time series
plot.TS_Overall(xts.c, type = "CO Concentrations",folder = "plots")
# Plot Dataframe with fautly rows in red
plot.DF_faults(df.new,df.specs = "Carbon Monoxide", avg_time, folder = "plots")


```

- Concentrations without corresponding wind speed and wind direction is highlighted in red.

```{r, echo = TRUE}
df.new.clean <- df.new[complete.cases(df.new), ]

# df.new$desea <- df.new$conc - df.new$sea
  
df.new.1 <- df.new[df.new$wos==1, ] 
```

## Wind Directions after 5min Downsampling

```{r wind_directions, echo = TRUE, fig.cap= "Wind Directions after 5min Downsampling"}
xts.ds.w <- xts(df.new.1$wdir, order.by = refTime[1:length(df.new.1$wdir)])
plot(xts.ds.w, main = paste("Wind Directions Downsampled by ",as.character(avg_time*60),"min",sep=""))
```

## 5 minute Downsampling
```{r wind_directions_hist, echo = TRUE, fig.height=5}
hist(df.new$wdir, main = "Histogram of wind directions", xlab = "Angle w.r.t to North")

```

## Temperature

```{r}
plot(df.new$temp, type = 'l')
```


# Seasonality from August dataset

```{r}

df.sea.w <- getSeasonalData_August(1)
plot.All_seasonality(df.sea.w, type = "all")
```

```{r}

plot.All_seasonality(df.sea.w, type = "overview")
```

## Quantile plots
```{r}
plot.weekQuantiles_Aug(avg_time, type = "overview")

```


# Effect of Wind Speed

## Effect of wind Speed on Concentrations

```{r sea_conc_ws, echo=TRUE, fig.height=5}

# df.new <- get_ts_master_dataFrame(sample.wd, sample.c, sample.ws, avg_time)
# refTime <- as.POSIXct(rownames(df.new))
df.seasonal <- getSeasonality(df.new, avg_time)
# Plots the daily seasonality w.r.t wind speed and wind direction
plot.DF_Seasonal_c_ws(df.seasonal, "Seasonal_plot_test", avg_time, "plots")

```

- With seasonality, the concentration observed has no bearing on the wind speed as observed from 18:00 to 00:00 hours.


## Correlation of concentration with wind speed
```{r ws_conc_corr, echo = TRUE, fig.height=5}
folder = "plots"
df.specs = "Corr_ws_c"
plot.Corr_ws_c(df.seasonal, df.specs, avg_time)
```

## Over view of all data

```{r}
folder = "plots"
df.specs = "dataset"
plot.dataset(df.new, df.seasonal, df.specs, avg_time)
```


## Effect of wind speed on de-seasoned data

```{r scat_ws_conc, echo = TRUE, warning= FALSE, fig.height= 5}
folder = "plots"
df.specs = "Scat_ws_deseasoned"
plot.Scat_ws_deseasoned(df.new, df.seasonal, df.specs, avg_time)

```

# Effect of wind direction on concentration

## Concentration vs Wind Direction

```{r warning=FALSE, echo = TRUE, fig.height= 5}
plot.Hist_conc_wd(df.new, df.specs, avg_time)
```


## Concentration vs Wind Direction

```{r avg_ws_wdir, echo  = TRUE, warning=FALSE, fig.height= 5}
# df.new.clean <- df.new[complete.cases(df.new), ]
tmp <- avg_time
avg_time <- 1/6
df.specs = "conc_ws_on_wd"
df.new <- getMasterdf_September(avg_time)
plot.conc_ws_on_wd(df.new, df.specs, avg_time)

avg_time = 1

# df.new <- get_ts_master_dataFrame('Raw Sensor Data/Wind_direction_corr_IISc_1_20.csv', 
#                               'Raw Sensor Data/Climo_co_corr_IISc_1_10.csv', 
#                               'Raw Sensor Data/Wind_speed_corr_IISc_1_20.csv', 
#                               avg_time)
df.new <- getMasterdf_September(avg_time)

```

- Wind directions $90^{\circ}$ and $315^{\circ}$ result in similar concentrations despite having higher wind speeds.

## Effect of Wind Direction on de-seasoned data

```{r scat_desea_ws_wdir, echo = TRUE}
# avg_time = 1

df.seasonal <- getSeasonality(df.new, avg_time)

plot.scat_ws_decon(df.new, df.seasonal, df.specs, avg_time)

```

# Analysis with Variance and Concentration

## Noise variance and concentration characteristic for Climo

```{r noise_var_conc_char, echo = TRUE, fig.height= 5, warning=FALSE, fig.cap="\\label{fig:var_C_C}"}
tmp <- avg_time
avg_time <- 1/6
df.specs = "NV_conc_char"
df.new <- getMasterdf_September(avg_time)

plot.NV_conc_char(df.new, df.seasonal, df.specs, avg_time)

avg_time = 1

# df.new <- get_ts_master_dataFrame('Raw Sensor Data/Wind_direction_corr_IISc_1_20.csv', 
#                               'Raw Sensor Data/Climo_co_corr_IISc_1_10.csv', 
#                               'Raw Sensor Data/Wind_speed_corr_IISc_1_20.csv', 
#                               avg_time)
df.new <- getMasterdf_September(avg_time)
df.seasonal <- getSeasonality(df.new, avg_time)



```


## Average of deseasoned concentrations conditioned by wind speed

```{r binning_ws_conc, echo = TRUE, fig.height=5, warning =FALSE}
df.specs = "conc_on_ws"
plot.conc_on_ws(df.new, df.seasonal, df.specs, avg_time)
```


## Z-score(Conc) vs Z-score(Wind speed) from 9AM to 6PM

```{r zscore_comp, echo = TRUE, fig.height=5, warning=FALSE, fig.cap="\\label{fig:z_W_C_var}"}
plot.Zconc_Zws(df.new, df.specs, avg_time)

```

# Deseasoning with weekly seasonality 

```{r, echo=TRUE, warning=FALSE}
# avg_time= 1/12
# 
# df.new <- getMasterdf_September(avg_time)

df.sea.d <- getSeasonality(df.new, avg_time)
df.sea.w <- getSeasonalData_August(avg_time)

plot.scat_ws_wdecon(df.new, df.sea.d,df.sea.w, avg_time)
```

# Effect of Temperatute



##  Seasonaltiy with concentration and temperature 
```{r}

plot.DF_Seasonal_c_t(df.seasonal, "Seasonal_plot_test", avg_time, "plots")

```


## Correlation of concentration with temperature

```{r t_conc_corr, echo = TRUE, fig.height=5}
folder = "plots"
df.specs = "Corr_t_c"
plot.Corr_t_c(df.seasonal, df.specs, avg_time)
```


## Effect of temperature on de-seasoned data

```{r scat_t_conc, echo = TRUE, warning= FALSE, fig.height= 5}
df.specs = "Scat_t_deseasoned"
plot.Scat_t_deseasoned(df.new, df.seasonal, df.specs, avg_time)

```
